---
title: "sorvi Tutorial"
author: rOpenGov core team
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{sorvi Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Finnish open (government) data toolkit for R
===========

## Introduction

sorvi package was originally intended for hosting various algorithms for Finnish open goverment data in [R](https://www.r-project.org). After being superseded by later rOpenGov packages that are more specialized in scope (such as [geofi](https://github.com/rOpenGov/geofi), [hetu](https://github.com/rOpenGov/hetu) and [pxweb](https://github.com/rOpenGov/pxweb)), sorvi has now found a new life in hosting various functions that are helpful in rOpenGov package maintenance, authoring and preparing presentations.

## Installation

Install the stable release version from CRAN in R:

```{r install, eval=FALSE}
install.packages("sorvi")
```

Alternatively, use `remotes` package to install the latest development version from GitHub:

```{r install2, eval=FALSE}
library(remotes)
remotes::install_github("ropengov/sorvi")
```

Loading the package in R command line:

```{r test, message=FALSE, warning=FALSE, eval=TRUE}
library(sorvi)
```

## Using the package

The number of functions in this package has been reduced compared to older versions. 

Get download statistics of eurostat-package, by year:

```{r example_download_stats, eval=FALSE}
df <- cran_downloads(pkgs = "eurostat", sum = "by_year", use.cache = FALSE)
df
```

Get download statistics of various rOpenGov packages over time and draw a chart:

```{r example_visualize, eval=FALSE}
packages <- c("eurostat", "giscoR", "sotkanet", "geofi", "sweidnumbr")
plot <- cran_downloads(pkgs = packages, sum = "total", output = "plot", use.cache = FALSE)
plot
```

## Licensing and Citations

This work can be freely used, modified and distributed under the 
[Two-clause BSD license](https://en.wikipedia.org/wiki/BSD\_licenses).

```{r citation, message=FALSE, eval=TRUE}
citation("sorvi")
```

## Session info

This vignette was created with

```{r sessioninfo, message=FALSE, warning=FALSE}
sessionInfo()
```

To call in the statistician after the experiment is done may be no more than asking him to perform a post-mortem examination: he may be able to say what the experiment died of. ~ Sir Ronald Aylmer Fisher
